Boosted

### 2015-10-23 boost_tree.rmd
finishing boosted

```{r, load library and packages, include= FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')

@@

all8a$b.rpdol<- all8a$return.pdol 
all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)
all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)

@@

@@

client.lim <- c(20,30,40,60,70)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
auc<- double(5)
colours <- c("red", "yellow", "green", "blue", "purple")

for(i in 1:5){

@@

        #redo train and test
        set.seed = 200
        sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
        train<- all9a[sample,]
        test<- all9a[-sample,]

        #make boosted tree
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 4000,
                 shrinkage=0.001,interaction.depth=3,
                 n.minobsinnode = 10)
        pred<- predict(fit, test, n.tree=4000, type="response")

@@

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")

}

@@

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[1:5])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

#best I can do is 0.783
```

Make a function to do the same with code.contact, JD.Second, and Business????

>>>>

```{r, load library and packages, include= FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('pROC',lib='C:/Progra~1/R/R-3.2.2/library')

@@

all8a$b.rpdol<- all8a$return.pdol 
all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)
all8a$pc.pro<- ifelse(all8a$pc.pro == 0 & is.na(all8a$pc.majpos.log), NA, all8a$pc.pro)
all8a$timespan.cbrt <- ifelse(all8a$timespan.cbrt ==0, 1, all8a$timespan.cbrt)

@@

Attempt at caret package with gbm

# must turn outcome into two level factor
all9a<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')[,-1]
all9a$f.rpdol<- as.factor(all9a$b.rpdol)
set.seed(100)
sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
train<- all9a[sample,]
test<- all9a[-sample,]

# set seeds
set.seed(123)
seeds <- vector(mode = "list", length = 51)
for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
## For the last model:
seeds[[51]] <- sample.int(1000, 1)

# first tune shrinkage and number of trees
train2<- train
train2$b.rpdol<- as.factor(train2$b.rpdol)
levels(train2$b.rpdol)[levels(train2$b.rpdol)=="0"] <- "profit"
levels(train2$b.rpdol)[levels(train2$b.rpdol)=="1"] <- "loss"

formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log +
client.totinv.log + pc.majpos.log + Billing.Type + majority.pos"

cv.Control <- trainControl(method = "cv",
                           number = 10,
                           seeds= seeds
                           , summaryFunction=twoClassSummary,
                           classProbs=TRUE
                           )
gbmGrid <- expand.grid(shrinkage = c(0.001),
                       n.trees = c(10000), interaction.depth = c(1,2,3,4,5),
                       n.minobsinnode = c(10,20,30))
set.seed(2)
gbmFit <- train(as.formula(formula), data= train2,
                 method = "gbm", trControl = cv.Control, verbose = FALSE,
                 bag.fraction = 0.5, tuneGrid = gbmGrid,
                metric = 'ROC',
                na.action = na.pass)

plot(gbmFit)
plot(gbmFit, plotType = 'level')

#conclusion:
# min num in terminal node = 10
# max tree depth = 4
# n.trees = 10000
# shrinkage = 0.001

After caret tuning, find AUC for test data set for set.seed = 100

```{r, echo = FALSE}
set.seed(300)
sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
train<- all9a[sample,]
test<- all9a[-sample,]

formula<- "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + pc.majpos.log + Billing.Type + majority.pos"

fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=4,
                 n.minobsinnode = 10)
pred<- predict(fit, test, n.tree=10000, type="response")

#pROC
pred.p<- roc(test$b.rpdol, pred)
plot.roc(pred.p, col= colours[i], print.thres = T)

#auc = 0.7788

@@

formula<- "b.rpdol~.- mlsto - return.pdol- Year- code.client - code.contact- JD.Second - Business-timespan.cbrt"

client.lim <- c(20,30,40,60,70, 90)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)), '6' = data.frame(c(0,0), c(0,0)))
auc<- rep(list(rep(0,6)),6)
colours <- c("red", "yellow", "green", "blue", "purple", "pink")

#set up train and test
        set.seed = 100
        sample<- sample(1:nrow(all9a), .75*nrow(all9a), replace=F)
        train<- all9a[sample,]
        test<- all9a[-sample,]
        #set up 10 fold cv using caret function
        folds <- createFolds(train$b.rpdol, k=10, list=T, returnTrain=F)
        #now have folds[[1]] through folds[[5]] list of rows to exclude

for(i in 1:6){

@@

        for (j in 1:10){
                #make boosted tree
                fit<-gbm(as.formula(formula), 
                         data=train[-folds[[j]],], distribution="bernoulli", n.trees= 10000,
                         shrinkage=0.001,interaction.depth=4,
                         n.minobsinnode = 10)

                pred<- predict(fit, train[folds[[j]], ], n.tree=10000, type="response")

                        #pROC
                        pred.p<- roc(train[folds[[j]],'b.rpdol'], pred)

                        if(i==1 & j==1){
                                plot.roc(pred.p, col= colours[i], print.thres = F)

                        } else{
                                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = F)
                        }

                        auc[[i]][j]<- auc(pred.p)

                        #confusion matrix
#                         conf.mat<- coords(pred.p, "best")
#                         pred1<- ifelse(pred>= conf.mat[1], 1, 0)
#                         temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
#                         conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
#                         cat(i," ")  
                        cat(j, " ")
        }

        print(i)        
}

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[[1:5]][6])
means<- sapply(auc, function(x) tail(x,1))
ggplot(shrink, aes(x=shrinkage, y=means)) + geom_line() + geom_point()
a<- ggplot(melt(auc), aes(x= factor(L1), y= value)) + geom_boxplot()

#save AUC for each 10 fold cross validation iteration, where client count cut off varied from 
client.cutoff<- melt(auc)
client.cutoff$L1<- as.factor(client.cutoff$L1)
levels(client.cutoff$L1)<- c("20","30","40","60","70", "90")
colnames(client.cutoff)[names(client.cutoff) %in% 'L1']<-'min.client.count'
colnames(client.cutoff)[names(client.cutoff) %in% 'value']<-'AUC'
write.csv(client.cutoff, 'C:/Users/n9232371/Documents/github/consult/finalwriteup/images/boosted tree/client_cutoff.csv')

a<- ggplot(client.cutoff, aes(x= min.client.count, y= AUC)) + geom_boxplot()
ggsave('C:/Users/n9232371/Documents/github/consult/finalwriteup/images/boosted tree/clientcode.jpg', a)

#client id useless for boosted trees!!

What about JD.Second?

```{r,echo=FALSE}

#compact JD.Second into less categories
all9a$JD.Second<- as.character(all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("heritage|expert", all9a$JD.Second), 
                          "heritage_expert", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("wharf|bridges", all9a$JD.Second), 
                          "wharf_bridge", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("membrane|art|awning|facade", all9a$JD.Second), 
                          "art_facade_awn_memb", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("hosp|emerg|health|carpark", all9a$JD.Second), 
                          "hosp_health_carpark", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("sign|product", all9a$JD.Second), 
                          "sign_product", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("hotel|office|dining", all9a$JD.Second), 
                          "hotel_office_dining", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("report|investigation", all9a$JD.Second), 
                          "report", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("waste water|management", all9a$JD.Second), 
                          "waste_wat_manage", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("univ|commun|school|exten", all9a$JD.Second), 
                          "edu_exten_community", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("multi|house", all9a$JD.Second), 
                          "residential", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("flood|harvest|recycled", all9a$JD.Second), 
                          "flood_h2o_harvest", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("bldg|subdiv|sewer", all9a$JD.Second), 
                          "civBldg_subdiv_sewer", all9a$JD.Second)
all9a$JD.Second <- ifelse(grepl("aged|erosion", all9a$JD.Second), 
                          NA, all9a$JD.Second)

all9a$JD.Second<- as.factor(all9a$JD.Second)

#now see if improvement with JD.Second

var.seq<-c("no.users + JD.Second", 
           "no.users")

conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)))
auc<- double(2)
colours <- c("red", "yellow")

#make fresh random sample of data
set.seed(1000)
sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
train<- all9a[sample,]
test<- all9a[-sample,]

for(i in 1:2){

        formula<- paste("b.rpdol~ Discipline + pc.pro + inv.mlsto.log + pc.majpos.log + majority.pos + b.timespan.cbrt + client.totinv.log + Billing.Type", var.seq[i], sep= " + ")

        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=4,
                 n.minobsinnode = 10)
        pred<- predict(fit, test, n.tree=10000, type="response")

@@

@@

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=1:2, "auc" = auc[1:2])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

# generally slightly better without JD.Second :(

What about Business?
